---
title: "Labs: title, subtitle & axis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{labs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Packages and data used below:

```{r message=FALSE, warning=FALSE}
library(apexcharter)
library(dplyr)

data("diamonds", package = "ggplot2")
n_cut <- count(diamonds, cut)
```


## Labs

You can set title, subtitle and axis' titles at once with `ax_labs()`:

```{r}
apex(data = n_cut, type = "column", mapping = aes(x = cut, y = n)) %>%
  ax_labs(
    title = "Cut distribution",
    subtitle = "Data from ggplot2",
    x = "Cut",
    y = "Count"
  )
```

If you more control (font size, alignment, ...), you can use `ax_title()`, `ax_subtitle()`, `ax_xaxis()` and `ax_yaxis()`, as described below.



## Title

```{r}
apex(data = n_cut, type = "column", mapping = aes(x = cut, y = n)) %>% 
  ax_title(text = "Cut distribution")
```

You can set some options, for example:

```{r}
apex(data = n_cut, type = "column", mapping = aes(x = cut, y = n)) %>% 
  ax_title(
    text = "Cut distribution", 
    align = "center",
    style = list(fontSize = "22px", fontWeight = 700)
  )
```


Full list of parameters is available here : https://apexcharts.com/docs/options/title/


## Subtitle

```{r}
apex(data = n_cut, type = "column", mapping = aes(x = cut, y = n)) %>% 
  ax_title(text = "Cut distribution") %>% 
  ax_subtitle(text = "Data from ggplot2")
```

With same options than for title:

```{r}
apex(data = n_cut, type = "column", mapping = aes(x = cut, y = n)) %>% 
  ax_title(
    text = "Cut distribution", 
    align = "center",
    style = list(fontSize = "22px", fontWeight = 700)
  ) %>% 
  ax_subtitle(
    text = "Data from ggplot2", 
    align = "center",
    style = list(fontSize = "16px", fontWeight = 400, color = "#BDBDBD")
  )
```

Full list of parameters is available here : https://apexcharts.com/docs/options/subtitle/


## Axis title

```{r}
apex(data = n_cut, type = "column", mapping = aes(x = cut, y = n)) %>% 
  ax_yaxis(title = list(text = "Count")) %>% 
  ax_xaxis(title = list(text = "Cut"))
```

With some options:

```{r}
apex(data = n_cut, type = "column", mapping = aes(x = cut, y = n)) %>% 
  ax_yaxis(title = list(
    text = "Count",
    style = list(fontSize = "14px", color = "#BDBDBD")
  )) %>% 
  ax_xaxis(title = list(
    text = "Cut", 
    style = list(fontSize = "14px", color = "#BDBDBD")
  ))
```




