---
title: "Customize chart options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customize chart options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Title, subtitle and axis titles

Packages and data used below:

```{r message=FALSE, warning=FALSE}
library(apexcharter)

data("diamonds", package = "ggplot2")
```


### Labs

You can set title, subtitle and axis' titles at once with `ax_labs()`:

```{r}
apex(data = diamonds, type = "column", mapping = aes(x = cut)) %>%
  ax_labs(
    title = "Cut distribution",
    subtitle = "Data from ggplot2",
    x = "Cut",
    y = "Count"
  )
```

If you want more control (font size, alignment, ...), you can use `ax_title()`, `ax_subtitle()`, `ax_xaxis()` and `ax_yaxis()`, as described below.



### Title

```{r}
apex(data = diamonds, type = "column", mapping = aes(x = cut)) %>% 
  ax_title(text = "Cut distribution")
```

You can set some options, for example:

```{r}
apex(data = diamonds, type = "column", mapping = aes(x = cut)) %>% 
  ax_title(
    text = "Cut distribution", 
    align = "center",
    style = list(fontSize = "22px", fontWeight = 700)
  )
```


Full list of parameters is available here : https://apexcharts.com/docs/options/title/


### Subtitle

```{r}
apex(data = diamonds, type = "column", mapping = aes(x = cut)) %>% 
  ax_title(text = "Cut distribution") %>% 
  ax_subtitle(text = "Data from ggplot2")
```

With same options than for title:

```{r}
apex(data = diamonds, type = "column", mapping = aes(x = cut)) %>% 
  ax_title(
    text = "Cut distribution", 
    align = "center",
    style = list(fontSize = "22px", fontWeight = 700)
  ) %>% 
  ax_subtitle(
    text = "Data from ggplot2", 
    align = "center",
    style = list(fontSize = "16px", fontWeight = 400, color = "#BDBDBD")
  )
```

Full list of parameters is available here : https://apexcharts.com/docs/options/subtitle/


### Axis title

```{r}
apex(data = diamonds, type = "column", mapping = aes(x = cut)) %>% 
  ax_yaxis(title = list(text = "Count")) %>% 
  ax_xaxis(title = list(text = "Cut"))
```

With some options:

```{r}
apex(data = diamonds, type = "column", mapping = aes(x = cut)) %>% 
  ax_yaxis(title = list(
    text = "Count",
    style = list(fontSize = "14px", color = "#BDBDBD")
  )) %>% 
  ax_xaxis(title = list(
    text = "Cut", 
    style = list(fontSize = "14px", color = "#BDBDBD")
  ))
```





## Lines

```{r message=FALSE, warning=FALSE}
library(apexcharter)

## economics dataset from ggplot2
data("economics", package = "ggplot2")
data("economics_long", package = "ggplot2")
economics <- tail(economics, 50)
economics_long <- subset(economics_long, date >= "2010-01-01")
```


### Type of line

Classic line:

```{r}
apex(data = economics, type = "line", mapping = aes(x = date, y = uempmed))
```

Spline curve:

```{r}
apex(data = economics, type = "line", mapping = aes(x = date, y = uempmed)) %>% 
  ax_stroke(curve = "smooth")
```

Steps chart:

```{r}
apex(data = economics, type = "line", mapping = aes(x = date, y = uempmed)) %>% 
  ax_stroke(curve = "stepline")
```



### Line appearance

Color line with gradient:

```{r}
apex(data = economics, type = "line", mapping = aes(x = date, y = uempmed)) %>% 
  ax_fill(
    type = "gradient",
    gradient = list(
      shade = "dark",
      gradientToColors = list("#FDD835"),
      shadeIntensity = 1,
      type = "horizontal",
      opacityFrom = 1,
      opacityTo = 1,
      stops = c(0, 100, 100, 100)
    )
  )
```

Solid area color:

```{r}
apex(data = economics, type = "area", mapping = aes(x = date, y = uempmed)) %>% 
  ax_fill(type = "solid", opacity = 1)
```


Line width:

```{r}
apex(data = economics, type = "line", mapping = aes(x = date, y = uempmed)) %>% 
  ax_stroke(width = 1)
```


Dotted line

```{r}
apex(data = economics, type = "line", mapping = aes(x = date, y = uempmed)) %>% 
  ax_stroke(dashArray = 6)
```



### Markers

Add points to line :

```{r}
apex(data = tail(economics, 20), type = "line", mapping = aes(x = date, y = uempmed)) %>% 
  ax_markers(size = 6)
```


Add labels over points

```{r}
apex(data = tail(economics, 20), type = "line", mapping = aes(x = date, y = uempmed)) %>% 
  ax_markers(size = 6) %>% 
  ax_dataLabels(enabled = TRUE)
```



### Multiple lines

You can use vectors of parameters to custom series separately:

```{r}
apex(data = economics_long, type = "line", mapping = aes(x = date, y = value01, group = variable)) %>% 
  ax_yaxis(decimalsInFloat = 2) %>% 
  ax_markers(size = c(3, 6)) %>% 
  ax_stroke(width = c(1, 3))
```


```{r}
apex(data = economics_long, type = "line", mapping = aes(x = date, y = value01, group = variable)) %>% 
  ax_yaxis(decimalsInFloat = 2) %>% 
  ax_stroke(dashArray = c(8, 5))
```

